Equivalence Hypergraphs: E-Graphs for Monoidal Theories

### 2024-01-11 E-Hypergraphs-LICS24.tex
Figures update up to Section 5

% \usepackage{mathtools,mathrsfs}
\usepackage{tikz-cd}
\usepackage{tikzit}

@@

The intuitive notion of (hyper-)graph rewriting is  as expected: given a rewrite rule $\mathcal L \rightsquigarrow \mathcal R$ and a graph $\mathcal{G}$,  we expect to find an occurrence of $\mathcal L$ within $\mathcal{G}$,  remove it,  and then plug $\mathcal R$ into the hole to recover the rewritten graph $\mathcal{H}$.  In DPO rewriting,  such a rewrite rule is formalized as a span $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$ in some appropriate category of graphs,  where $\mathcal K$ is some invariant that should be maintained during rewriting.
A typical DPO square looks as follows,  where the two marked squares are pushouts. 
\[
\begin{tikzcd}
\mathcal L \arrow[d, "m"'] & \mathcal K \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
\mathcal G^{\urcorner}     & \mathcal C \arrow[l] \arrow[r]                      & ^{\ulcorner} \mathcal H
\end{tikzcd}
\]

@@

To express the rewriting of (hyper-)graphs with interfaces,  \textit{i.e.},  the ones that represent morphisms in $\HypI{\Sigma}$, the DPO formalism has been extended to double-pushout rewriting with \textit{interfaces}  (DPOI).
In the case of $ \catname{Hyp(\Sigma)}$,  we wish rewrite rules to be pairs of discrete cospans of hypergraphs with matching interfaces: $n \xrightarrow{} \mathcal L \xleftarrow{} m$, $n \xrightarrow{} \mathcal R \xleftarrow{} m$.  By noting that a cospan $0 \xrightarrow{} \mathcal L \xleftarrow{[f,g]} n + m$ encodes the same data as a cospan $n \xrightarrow{f} \mathcal L \xleftarrow{g} m$ (Remark 3.14~\cite{bonchi_string_2022-1}) we can formulate a rewrite rule as a single span $\mathcal L \xleftarrow{} n+m \xrightarrow{} \mathcal R$,  and we can do similarly for the discrete cospans giving the interface to $\mathcal G$ and $\mathcal H$.  Thus we reach the following variant of DPO rewriting. 
\begin{definition}[DPOI rewriting]\label{def:dpoi}
Given a span of morphisms $\mathcal G \xleftarrow{} n+m \xrightarrow{} \mathcal H$ in $\textbf{Hyp}(\Sigma)$,  we say \textit{$\mathcal G$ rewrites to $\mathcal H$ with interface $n+m$ via rewrite rule $\mathcal L \xleftarrow{} i+j \xrightarrow{} \mathcal R$} if there exists an object $\mathcal C$ and morphisms which complete the following commutative diagram such that the two marked squares are pushouts.
\[
\begin{tikzcd}
\mathcal L \arrow[d, "m"'] & i+j \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
\mathcal{G}^{\urcorner}     & \mathcal C \arrow[l] \arrow[r]                      & ^{\ulcorner}\mathcal H \\
                  & n+m \arrow[lu] \arrow[u] \arrow[ru]          &              
\end{tikzcd}
\]
\end{definition}
We need to restrict the notions of pushout complement and of match in order to preserve monogamy and directed acylcicity.

@@

% https://q.uiver.app/#q=WzAsNSxbMCwwLCJMIl0sWzIsMCwiaStqIl0sWzAsMiwiR157XFx1cmNvcm5lcn0iXSxbMiw0LCJuK20iXSxbMiwyLCJMXntcXGJvdH0iXSxbMyw0LCJbZF8xLGRfMl0iLDJdLFszLDIsIltiXzEsYl8yXSJdLFswLDIsImYiLDJdLFsxLDAsImEgPSBbYV8xLGFfMl0iLDJdLFsxLDQsImMgPSBbY18xLGNfMl0iXSxbNCwyLCJnIl1d
\[\begin{tikzcd}
	\mathcal L && {i+j} \\
	\\
	{\mathcal G^{\urcorner}} && {\mathcal{L}^{\bot}} \\
	\\
	&& {n+m}
	\arrow["{[d_1,d_2]}"', from=5-3, to=3-3]
	\arrow["{[b_1,b_2]}", from=5-3, to=3-1]
	\arrow["f"', from=1-1, to=3-1]
	\arrow["{[a_1,a_2]}"', from=1-3, to=1-1]
	\arrow["{[c_1,c_2]}", from=1-3, to=3-3]
	\arrow["g", from=3-3, to=3-1]
\end{tikzcd}\]
is called a boundary complement if $[c_1, c_2]$ is mono and there exist $d_1 : n \to \mathcal{L}^{\bot}$ and $d_2 : m \to \mathcal{L}^{\bot}$ making the above triangle commute and such that
\[
    n + j \xrightarrow{[d_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[d_2,c_1]} m + i
\]
is a monogamous cospan. 
\end{definition}
%\begin{remark}
%    Note how the requirement for $i \xrightarrow{} L \xleftarrow{} j$ and $n \xrightarrow{} G \xleftarrow{} m$ being monogamous, i.e. morphisms in $\MdaCospans$, is a part of the above definition.
%\end{remark}
Intuitively,  requiring boundary complements ensures that inputs are glued exclusively to outputs and vice-versa in the two pushout squares.

@@

    Two cospans are isomorphic if there exist isomorphisms $\alpha$, $\beta$ and $\gamma$ making the following diagram commute.
    \[
    \scalebox{0.75}{
        \tikzfig{combinatorial_semantics/isomorphic_e_cospans}
    }
    \]
Composition of two morphisms $f : X_{ext} \to Y_{ext}$ and $g : Y_{ext} \to Z_{ext}$ is computed in two stages
First, $\mathcal{F} +_{f_{ext}, g_{ext}} \mathcal{G}$ is computed,  as shown below.

@@

\[
\scalebox{0.9}
{\tikzfig{combinatorial_semantics/pushout_e_cospans}}
\]
Then, new feet for $\mathcal{H} = A +_{f_{ext};f,g_{ext};g} B$ are computed as below
\[
X_{ext} \xrightarrow{a_{ext};\iota_1} X \uplus (Y' \setminus Y_{ext}) \xrightarrow{h} \mathcal{H} \xleftarrow{k} Z \uplus (Y \setminus Y_{ext}) \xleftarrow{g_{ext};\iota_1} Z_{ext}
\]
where
\[
    h = [ a;p_1, ~g';p_2 ]
\qquad
    k = [ b;p_2, ~f';p_1 ]
\]
% \[
% \begin{tikzcd}
% Y_{ext} \arrow[d] & 0 \arrow[l] \arrow[d]    \\
% Y'          & Y' \setminus Y_{ext} \arrow[l]
% \end{tikzcd}
% \]
% and then the disjoint union is just a coproduct. 
where $g'$ is the restriction of $g$ to ${Y' \setminus Y_{ext}}$ and $f'$ the restriction of $f$ to $Y \setminus Y_{ext}$. 
The identity of composition is given by the obvious extended cospan of identities. 
$\Ecospans$ inherits a symmetric monoidal structure from the coproduct (and initial object) structure of $\catname{EHyp({\Sigma}})$,  analogously to Definition~\ref{def:cspd}.  
\end{definition}
CHRIS: UPDATE DIAGRAMS TO USE MY NEW NOTATION,  TAKE LESS VERTICAL SPACE.  FIND CLEANER WAY TO WRITE NEW FEET.  USE $\iota$ FOR INJECTIONS

@@

CHRIS: REMARK THAT WE CANNOT DEAL WITH SIGNATURES WITH $f:0 \to 0$ MORPHISMS, AND WHY!
Because extended cospans have a more general notion of interface, DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics} needs some adjustments.
First, the existence of \textit{internal} interfaces needs to be accounted for,  and in particular that internal interfaces are not necessarily preserved during rewriting.
The former means that rewrite rules become of the form
\[
\mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
\] where we use the fact that cospans 
\[
    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
\]
and
\[
    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j
\]
encode the same data and we will further omit the arrows $0 \to 0 \to \mathcal{L}$.
That is, a rewrite rule is a pair of cospans of e-hypergraphs with matching \textit{external} interfaces and arbitrary \textit{internal} interfaces.
To illustrate the non-preservation of internal interfaces, consider the rewrite rule above and the following e-hypergraph with interfaces to perform the rewrite on
% \[\langle x \xrightarrow{} x' \xrightarrow{} \mathcal{L} \xleftarrow{} y' \xleftarrow{} y, x \xrightarrow{} x'' \xrightarrow{} \mathcal{R} \xleftarrow{} y'' \xleftarrow{} y \rangle \]
% such that $x' \not \cong x''$ and $y' \not \cong y''$ and consider the following e-hypergraph to perform the rewrite on
\[
    \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k
\] such that there is a matching $m : \mathcal{L} \to \mathcal{G}$  and assume $i' \not \cong i''$ and $j' \not \cong j''$.
When removing the occurence of $\mathcal{L}$ from $\mathcal{G}$, \emph{i.e.}, by computing the pushout complement, the internal interfaces of $\mathcal{G}$ should be modified to exclude the vertices that were a part of (the image of) $\mathcal{L}$'s strict internal interfaces.
Then, when gluing $\mathcal{R}$ into the created context, we need to introduce strict internal interfaces of $\mathcal{R}$ into this context.
In particular, it means that the internal interfaces of the resulting cospan of e-hypergraphs
\[
    \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k 
\]
not necessarily coincides with the internal interfaces of the initial cospan.
% The input interface on the left-hand side of the rule is $\{1, 2\}$, while the input interface on the right-hand side of the rule is $\{1,2,5,6,7,8\}$.
% Note that the outermost interfaces are preserved.

@@

For monogamous extended cospans 
\[
    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
\quad\text{ and }\quad
    n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
\] and mono $f : \mathcal{L} \to \mathcal{G}$, a pushout complement 
\[
    i + j \to \mathcal{L}^{\bot} \to G^{\urcorner}
\] 
as depicted in the square below
\[
\scalebox{0.75}{

@@

is a \textit{boundary} complement if $[c_1,c_2]$ is mono 
 and there exist $d_1 : n \to C$ and $d_2 : k \to C$ 
 making the above triangle commute and such that
 DAN: THIS (AND SUCH THAT) IS EXTREMELY UNCLEAR, PLEASE REWORD
there exists a \textit{not necessarily well-typed} mda-cospan

\[
n \xrightarrow{} l + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} r + i \xleftarrow{} k
\]

where $l$ and $r$ are
\[
    l = n' \setminus (i' \setminus i)
\quad\text{ and }\quad
    r = k' \setminus (j' \setminus j).
\]
Because $\mathcal{L}^{\bot}$ is essentially $\mathcal{G}$ with $\mathcal{L}$ removed, apart from the part that has a pre-image in $i + j$, there is an obvious morphism from $l \to \mathcal{L}^{\bot}$ which is $g_1 = g_{int} |_{D}$ where
\[
    D = \{ v \in n' \text{ such that } \not \exists u \in (i' \setminus i) ~ . ~ g_{int}(v) = f'_{int};f(u) \}
\]
Similarly for $r \to \mathcal{L}^{\bot}$, which we denote as $g_2$.
Then there exists a morphism $n \to l$, because there is a morphism $n \to n'$ and the external interfaces of $n'$ and $l$ are the same.
Similarly for $k \to r$.
\end{definition}

@@

The conditions (3)-(4), in particular, prohibit a match $\{f \mapsto f, g \mapsto g\}$ in $\mathcal{L} = f \otimes g$ and $\mathcal{G} = f + g$.

>>>>

% \usepackage{mathtools,mathrsfs}
\usepackage{tikz-cd}
\usepackage{tikzit}
% to reduce tikz picture white space
\usepackage{adjustbox}

@@

The intuitive notion of (hyper-)graph rewriting is  as expected: given a rewrite rule $\mathcal L \rightsquigarrow \mathcal R$ and a graph $\mathcal{G}$,  we expect to find an occurrence of $\mathcal L$ within $\mathcal{G}$,  remove it,  and then plug $\mathcal R$ into the hole to recover the rewritten graph $\mathcal{H}$.  In DPO rewriting,  such a rewrite rule is formalized as a span $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$ in some appropriate category of graphs,  where $\mathcal K$ is some invariant that should be maintained during rewriting.
A typical DPO square looks as follows,  where the two marked squares are pushouts. 
% \[
% \begin{tikzcd}
% \mathcal L \arrow[d, "m"'] & \mathcal K \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
% \mathcal G^{\urcorner}     & \mathcal C \arrow[l] \arrow[r] & ^{\ulcorner} \mathcal H
% \end{tikzcd}
% \]
\[
    \scalebox{0.8}{
        \tikzfig{combinatorial_semantics/DPO_square}
    }
\]

@@

To express the rewriting of (hyper-)graphs with interfaces,  \textit{i.e.},  the ones that represent morphisms in $\HypI{\Sigma}$, the DPO formalism has been extended to double-pushout rewriting with \textit{interfaces}  (DPOI).
In the case of $ \catname{Hyp(\Sigma)}$,  we wish rewrite rules to be pairs of discrete cospans of hypergraphs with matching interfaces: $n \xrightarrow{} \mathcal L \xleftarrow{} m$, $n \xrightarrow{} \mathcal R \xleftarrow{} m$.  By noting that a cospan $0 \xrightarrow{} \mathcal L \xleftarrow{[f,g]} n + m$ encodes the same data as a cospan $n \xrightarrow{f} \mathcal L \xleftarrow{g} m$ (Remark 3.14~\cite{bonchi_string_2022-1}) we can formulate a rewrite rule as a single span $\mathcal L \xleftarrow{} n+m \xrightarrow{} \mathcal R$,  and we can do similarly for the discrete cospans giving the interface to $\mathcal G$ and $\mathcal H$.  Thus we reach the following variant of DPO rewriting. 
\begin{definition}[DPOI rewriting]\label{def:dpoi}
Given a span of morphisms $\mathcal G \xleftarrow{} n+m \xrightarrow{} \mathcal H$ in $\textbf{Hyp}(\Sigma)$,  we say \textit{$\mathcal G$ rewrites to $\mathcal H$ with interface $n+m$ via rewrite rule $\mathcal L \xleftarrow{} i+j \xrightarrow{} \mathcal R$} if there exists an object $\mathcal C$ and morphisms which complete the following commutative diagram such that the two marked squares are pushouts.
% \[
% \begin{tikzcd}
% \mathcal L \arrow[d, "m"'] & i+j \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
% \mathcal{G}^{\urcorner}     & \mathcal C \arrow[l] \arrow[r]                      & ^{\ulcorner}\mathcal H \\
%                   & n+m \arrow[lu] \arrow[u] \arrow[ru]          &              
% \end{tikzcd}
% \]
\[
    \scalebox{0.8}{
        \tikzfig{combinatorial_semantics/DPOI_square_Hyp}
    }
\]
\end{definition}
We need to restrict the notions of pushout complement and of match in order to preserve monogamy and directed acylcicity.

@@

% https://q.uiver.app/#q=WzAsNSxbMCwwLCJMIl0sWzIsMCwiaStqIl0sWzAsMiwiR157XFx1cmNvcm5lcn0iXSxbMiw0LCJuK20iXSxbMiwyLCJMXntcXGJvdH0iXSxbMyw0LCJbZF8xLGRfMl0iLDJdLFszLDIsIltiXzEsYl8yXSJdLFswLDIsImYiLDJdLFsxLDAsImEgPSBbYV8xLGFfMl0iLDJdLFsxLDQsImMgPSBbY18xLGNfMl0iXSxbNCwyLCJnIl1d
% \[\begin{tikzcd}
% 	\mathcal L && {i+j} \\
% 	\\
% 	{\mathcal G^{\urcorner}} && {\mathcal{L}^{\bot}} \\
% 	\\
% 	&& {n+m}
% 	\arrow["{[d_1,d_2]}"', from=5-3, to=3-3]
% 	\arrow["{[b_1,b_2]}", from=5-3, to=3-1]
% 	\arrow["f"', from=1-1, to=3-1]
% 	\arrow["{[a_1,a_2]}"', from=1-3, to=1-1]
% 	\arrow["{[c_1,c_2]}", from=1-3, to=3-3]
% 	\arrow[from=3-3, to=3-1]
% \end{tikzcd}\]

\[
    \scalebox{0.8}{
        \tikzfig{combinatorial_semantics/DPOI_boundary_complement}
    }
\]

is called a boundary complement if $[c_1, c_2]$ is mono and there exist $d_1 : n \to \mathcal{L}^{\bot}$ and $d_2 : m \to \mathcal{L}^{\bot}$ making the above triangle commute and such that
\[
    n + j \xrightarrow{[d_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[d_2,c_1]} m + i
\]
is a monogamous cospan. 
\end{definition}
%\begin{remark}
%    Note how the requirement for $i \xrightarrow{} L \xleftarrow{} j$ and $n \xrightarrow{} G \xleftarrow{} m$ being monogamous, i.e. morphisms in $\MdaCospans$, is a part of the above definition.
%\end{remark}
Intuitively,  requiring boundary complements ensures that inputs are glued exclusively to outputs and vice-versa in the two pushout squares.

@@

    Two cospans are isomorphic if there exist isomorphisms $\alpha$, $\beta$ and $\gamma$ making the following diagram commute.
    \[
    \scalebox{0.75}{
        \tikzfig{combinatorial_semantics/isomorphic_e_cospans}
    }
    \]
Composition of two morphisms $h_1 : n \to m$ and $h_2 : m \to k$ is computed in two stages.
First, $\mathcal{H} = \mathcal{F} +_{f_{ext}, g_{ext}} \mathcal{G}$ is computed,  as shown below.

@@

\[
\trimbox{0cm 0cm 0cm 1cm}{
\adjustbox{scale=0.9,center}
{\tikzfig{combinatorial_semantics/pushout_e_cospans}}
}
\]
Then, new feet for $\mathcal{H}$ are computed as below
\[
n \xrightarrow{f_{ext};\iota_1} n' \uplus (m'' \setminus m) \xrightarrow{h_{1}} \mathcal{H} \xleftarrow{h_2} k' \uplus (m' \setminus m) \xleftarrow{g'_{ext};\iota_1} k
\]
where
\[
    h_1 = [ f_{int};p_1, ~g_{int}|_{m'' \setminus m};p_2 ]
\qquad
    h_2 = [ g'_{int};p_2, ~f'_{int}|_{m' \setminus m};p_1 ]
\]
% \[
% \begin{tikzcd}
% Y_{ext} \arrow[d] & 0 \arrow[l] \arrow[d]    \\
% Y'          & Y' \setminus Y_{ext} \arrow[l]
% \end{tikzcd}
% \]
% and then the disjoint union is just a coproduct. 
% where $g'$ is the restriction of $g$ to ${Y' \setminus Y_{ext}}$ and $f'$ the restriction of $f$ to $Y \setminus Y_{ext}$. 
The identity of composition is given by the obvious extended cospan of identities. 
$\Ecospans$ inherits a symmetric monoidal structure from the coproduct (and initial object) structure of $\catname{EHyp({\Sigma}})$,  analogously to Definition~\ref{def:cspd}.  
\end{definition}
CHRIS: UPDATE DIAGRAMS TO USE MY NEW NOTATION,  TAKE LESS VERTICAL SPACE.  FIND CLEANER WAY TO WRITE NEW FEET.  USE $\iota$ FOR INJECTIONS

@@

CHRIS: REMARK THAT WE CANNOT DEAL WITH SIGNATURES WITH $f:0 \to 0$ MORPHISMS, AND WHY!
Because extended cospans have a more general notion of interface, DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics} needs some adjustments.
First, the existence of \textit{internal} interfaces needs to be accounted for,  and in particular that internal interfaces are not necessarily preserved during rewriting.
The former means that we wish a rewrite rule to be a pair of \textit{extended} cospans of e-hypergraphs with matching external interfaces.

\[
    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
\]
and
\[
    i \xrightarrow{} i'' \xrightarrow{} \mathcal{R} \xleftarrow{} j'' \xleftarrow{} j
\]

We will next use the fact that cospans
\[
    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
\]
and
\[
    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j
\]
endcode the same data and only consider cospans of the form

\[
\mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
\] to fit the DPO formalism.
We will futher omit the arrows $0 \to 0 \to \mathcal{L}$.
To illustrate the non-preservation of internal interfaces, consider the rewrite rule above and the following e-hypergraph with interfaces to perform the rewrite on
% \[\langle x \xrightarrow{} x' \xrightarrow{} \mathcal{L} \xleftarrow{} y' \xleftarrow{} y, x \xrightarrow{} x'' \xrightarrow{} \mathcal{R} \xleftarrow{} y'' \xleftarrow{} y \rangle \]
% such that $x' \not \cong x''$ and $y' \not \cong y''$ and consider the following e-hypergraph to perform the rewrite on
\[
    \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k
\] such that there is a matching $m : \mathcal{L} \to \mathcal{G}$  and assume $i' \not \cong i''$ and $j' \not \cong j''$.
When removing the occurence of $\mathcal{L}$ from $\mathcal{G}$, \emph{i.e.}, by computing the pushout complement, the internal interfaces of $\mathcal{G}$ should be modified to exclude the vertices that were a part of (the image of) $\mathcal{L}$'s strict internal interfaces.
Then, when gluing $\mathcal{R}$ into the created context, we need to introduce strict internal interfaces of $\mathcal{R}$ into this context.
In particular, it means that the internal interfaces of the resulting cospan of e-hypergraphs
\[
    \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k 
\]
not necessarily coincides with the internal interfaces of the initial cospan.
% The input interface on the left-hand side of the rule is $\{1, 2\}$, while the input interface on the right-hand side of the rule is $\{1,2,5,6,7,8\}$.
% Note that the outermost interfaces are preserved.

@@

For monogamous extended cospans 
\[
    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
\quad\text{ and }\quad
    n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
\] and mono $f : \mathcal{L} \to \mathcal{G}$, a pushout complement 
\[
    i + j \to \mathcal{L}^{\bot} \to \mathcal{G}
\] 
as depicted in the square below
\[
\scalebox{0.75}{

@@

is a \textit{boundary} complement if the following holds
\begin{enumerate}
    \item $[c_1,c_2]$ is mono 
    \item there exist $d_1 : n \to \mathcal{C}$ and $d_2 : k \to \mathcal{C}$ making the above triangle commute
    \item there exists a \textit{not necessarily well-typed} mda-cospan
    \[
    n \xrightarrow{} l + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} r + i \xleftarrow{} k
    \]
\end{enumerate}

where $l$ and $r$ are
\[
    l = n' \setminus (i' \setminus i)
\quad\text{ and }\quad
    r = k' \setminus (j' \setminus j).
\]
Because $\mathcal{L}^{\bot}$ is essentially $\mathcal{G}$ with $\mathcal{L}$ removed, apart from the part that has a pre-image in $i + j$, there is an obvious morphism from $l \to \mathcal{L}^{\bot}$ which is $g_1 = g_{int} |_{D}$ where
\[
    D = \{ v \in n' \text{ such that } \not \exists u \in (i' \setminus i) ~ . ~ g_{int}(v) = f'_{int};f(u) \}
\]
Similarly for $r \to \mathcal{L}^{\bot}$, which we denote as $g_2$.
Then there exists a morphism $n \to l$, because there is a morphism $n \to n'$ and the external interfaces of $n'$ and $l$ are the same.
That is, it is a morphism $f: n \to l$ such that $d_1 = f;g_1$.
Similarly for $k \to r$.
\end{definition}

@@

The conditions (3)-(4), in particular, prohibit a match $\{f \mapsto f, g \mapsto g\}$ in $\mathcal{L} =$ $\scalebox{0.4}{\tikzfig{combinatorial_semantics/f_times_g}}$ and $\mathcal{G} =$ $\scalebox{0.4}{\tikzfig{combinatorial_semantics/f_plus_g_inline}}$.
